Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2024-03-29 2_SSL_NE/article/article_main.tex
added section to supplement

\begin{abstract}
Since their introduction to the market in 1996, white light-emitting diodes (LEDs) have greatly improved in performance, efficiency, and manufacturing cost. These advancements have been crucial for reducing global carbon emissions in the lighting sector which accounts for nearly 20\% of global electricity consumption. Understanding the extent and sources of the rapid progress in white LED technology can thus provide valuable insights for accelerating innovation in other demand-side clean energy technologies. Here we show, through cost and performance modeling based on data from literature review, patent analysis, and expert interviews, that the efficiency of top-performing warm white GaN-based LED packages increased from 5.8\% in 2003 to 38.8\% in 2020. Over the same period, the manufacturing cost of low-to-mid-power LED packages decreased by 95.5\% from \$1.1 to \$0.05 (in 2020 USD). We find that technology spillovers from other sectors accounted for at least 8.5\% of efficiency improvements and nearly all consumer experience enhancements, playing an important role in widespread LED adoption in lighting.
\end{abstract}

>>>>

\begin{abstract}
\hl{
Since their introduction to the market in 1996, white light-emitting diodes (LEDs) have greatly improved in performance, efficiency, and manufacturing cost. Understanding the extent and mechanisms of the rapid progress in white LED technology can provide valuable insights for accelerating innovation in other demand-side clean energy technologies critical for reducing global carbon emissions. Here we show, through cost and performance modeling based on data from literature review, patent analysis, and expert interviews, that the efficiency of top-performing warm white GaN-based LED packages increased from 5.8\% in 2003 to 38.8\% in 2020. Over the same period, the manufacturing cost of low-to-mid-power LED packages decreased by 95.5\% from \$1.1 to \$0.05 (in 2020 USD). We find that technology spillovers from other sectors accounted for at least 8.5\% of efficiency improvements and nearly all consumer experience enhancements, playing an important role in widespread LED adoption in lighting.
}
\end{abstract}

### 2024-03-29 2_SSL_NE/article/article_supplementary.tex
added section to supplement

% custom section designations
% https://tex.stackexchange.com/a/712523

\newcounter{suppnotecounter}
\newcommand{\suppnote}[1]{%
  \stepcounter{suppnotecounter}%
  \subsection*{\large\bfseries\filright Supplementary Note \arabic{suppnotecounter}: \\ #1}%
  \addcontentsline{toc}{subsection}{#1}%
  \label{suppnote:#1}
}

\newcounter{suppdisccounter}
\newcommand{\suppdisc}[1]{%
  \stepcounter{suppdisccounter}%
  \subsection*{\large\bfseries\filright Supplementary Discussion \arabic{suppdisccounter}: \\ #1}%
  \addcontentsline{toc}{subsection}{#1}%
}

\newcounter{suppmethcounter}
\newcommand{\suppmeth}[1]{%
  \stepcounter{suppmethcounter}%
  \subsection*{\large\bfseries\filright Supplementary Method \arabic{suppmethcounter}: \\ #1}%
  \addcontentsline{toc}{subsection}{#1}%
}

\newcounter{supprefcounter}
\newcommand{\suppref}[1]{%
  \stepcounter{supprefcounter}%
  \subsection*{\large\bfseries\filright Supplementary References \arabic{supprefcounter}: \\ #1}%
  \addcontentsline{toc}{subsection}{#1}%
}

@@

\addcontentsline{toc}{section}{Supplementary Notes}

\suppnote{Previous Literature on Technological Progress in LEDs}
\label{sec:prev_lit}

\cref{sec:prev_lit}

@@

\suppnote{Manufacturing Process Steps}
\label{sec:manufacturing_steps}

@@

\suppnote{Historical Progress in Chip Architectures}
\label{sec:chip_architecture}

\begin{figure}[h!]
\centering
\includegraphics[height=15cm]{figures/chip_architecture_overview.pdf}
\caption{\textbf{Historical evolution of LED chip architectures.} Panel A: classical chip with lateral current spreading; Panel B: Osram’s thin GaN flip-chip (vertical) architecture; Panel C: flip-chip architecture; Panel D: chip-scale package flip-chip architecture. Shown are side views of chips without packages, along a cutaway line best suited to the features of each architecture. The cutaway surface is indicated by a red arrow with an eye on the overlaid top view of each chip. Note that the dimensions are not to scale, and smaller features are greatly exaggerated for clarity. Years indicated correspond to the earliest identified patent priority date. Black frames around certain designs indicate chip designs not brought to large-scale production. Chip architecture abbreviations: TF - Thin-Film; FC - Flip-Chip; CSP - Chip-Scale Package; Material abbreviations: GaN - Gallium Nitride, ITO - Indium Tin Oxide, MQW - Multiple Quantum Well. Source: adapted and compiled from multiple patents and industry publications. See Supplementary Note 16 for the full list of sources and references.}
\label{fgr:chip_architecture_overview}
\end{figure}

We summarize the information collected and analysed on the progress in GaN-based LED chip architectures and manufacturing processes in \cref{fgr:chip_architecture_overview}, where we show the evolution of LEDs from classical chips with lateral current spreading to chip-scale package flip-chip architectures. We use this information as input into LED performance calculations and manufacturing cost modeling.

\clearpage
\suppnote{Historical Progress in Sub-Efficiencies}
\label{sec:subeff_progress}

\begin{figure}[H]
\centering
\includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
\caption{\textbf{Historical improvements in LED sub-efficiencies.} Panels A1-A2 show data for key metrics used to calculate forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show data for internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Source: own synthesis of data from the list of sources provided in \cref{sec:sources}. For an illustration of the different chip architectures shown, compare \cref{sec:chip_architecture}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
\label{fgr:subeff}
\vspace{-20mm}
\end{figure}

@@

\textbf{Additional Notes on \cref{fgr:subeff}}

Data illustrated in \cref{fgr:subeff} was used to determine the performance of best-in-class devices across all sub-efficiencies. The results are shown in Figures 4-6 in the main text.

@@

\textbf{Additional Data on Phosphor Performance}

Note that \cref{fgr:subeff} does not include sub-efficiencies related to spectral performance of LEDs: spectral efficiency $\eta_S$ and conversion efficiencies of different phosphors $\eta_{C_R}$ and $\eta_{C_{Y/G}}$.  Typically, there is no direct empirical data available for these metrics in the literature, which made their inclusion in the (mostly) empirical \cref{fgr:subeff} not relevant. Instead, we calculate the values of these metrics used in Figures 4 and 5 in the main text from spectral and luminous efficacy radiation data presented in \cref{fig:phosphor_spectrum} below (see also Methods section in the main text).

\clearpage
\suppnote{Non-Phosphor-Related LED Innovations}

@@

\clearpage
\suppnote{Phosphor-Related LED Innovations}
\label{sec:innovation_phosphor}

@@

\suppnote{Spectral Data for Phosphors}

@@

\clearpage
\suppref{Full List of Sources for Figures}
\label{sec:sources}

@@

\clearpage
\addcontentsline{toc}{section}{Supplementary Methods}

\suppmeth{Domain Definition}
\label{sec:domain}

\textit{Elsevier Scopus}\footnote{\url{https://www.scopus.com/search/}} advanced search query:

\begin{pseudocode}
TITLE-ABS-KEY(("light-emitting" OR "light emitting") W/2 diode*)
AND NOT
TITLE-ABS-KEY(solar OR violet OR laser)
AND NOT
TITLE-ABS-KEY(perovskite OR GaAsP OR AlGaInP OR GaAlAs OR ZnO)
AND NOT
TITLE-ABS-KEY(organic OR OLED* OR polymer*)
AND NOT
SUBJAREA(ARTS OR BUSI OR DECI OR ECON OR PSYC OR SOCI)
AND NOT
SUBJAREA(MEDI OR NURS OR VETE OR DENT OR HEAL OR MULT)
AND NOT
SUBJAREA(AGRI OR BIOC OR IMMU OR NEUR OR PHAR)
AND NOT
SUBJAREA(envi OR EART OR comp)
\end{pseudocode}

@@

\suppmeth{Metrics for Quantifying Technological Progress in LEDs}

@@

\clearpage
\suppmeth{Device Performance Metrics}
\label{sec:device_performance_metrics}

@@

\clearpage
\suppmeth{Consumer Experience Metrics}
\label{sec:metrics_consumer}

@@

\suppmeth{Background of Interviewed Experts}

@@

\clearpage
\suppmeth{Cost Model Structure and Limitations}
\label{sec:costmodel_structure}

@@

\suppmeth{Yielded Cost}
\label{sec:cost_computation}

@@

\suppmeth{Contribution of Variables}
\label{sec:contribution_variables}

@@

\addcontentsline{toc}{section}{Supplementary Discussion}

\suppdisc{Cost Model Sensitivity Analysis}

@@

\clearpage
\suppdisc{Cost Data Comparison with Industry Projections}
\label{sec:cost_model_calibration}

>>>>

@@

\section{Supplementary Notes}

\subsection{Previous Literature on Technological Progress in LEDs}
\label{sec:prev_lit}

@@

\subsection{Manufacturing Process Steps}
\label{sec:manufacturing_steps}

@@

\subsection{Historical Progress in Chip Architectures}
\label{sec:chip_architecture}

\begin{figure}[h!]
\centering
\includegraphics[height=15cm]{figures/chip_architecture_overview.pdf}
\caption{\textbf{Historical evolution of LED chip architectures.} Panel A: classical chip with lateral current spreading; Panel B: Osram’s thin GaN flip-chip (vertical) architecture; Panel C: flip-chip architecture; Panel D: chip-scale package flip-chip architecture. Shown are side views of chips without packages, along a cutaway line best suited to the features of each architecture. The cutaway surface is indicated by a red arrow with an eye on the overlaid top view of each chip. Note that the dimensions are not to scale, and smaller features are greatly exaggerated for clarity. Years indicated correspond to the earliest identified patent priority date. Black frames around certain designs indicate chip designs not brought to large-scale production. Additional context on this figure is provided in \cref{sec:subeff_progress}. Chip architecture abbreviations: TF - Thin-Film; FC - Flip-Chip; CSP - Chip-Scale Package; Material abbreviations: GaN - Gallium Nitride, ITO - Indium Tin Oxide, MQW - Multiple Quantum Well. Source: adapted and compiled from multiple patents and industry publications. See Supplementary Note 16 for the full list of sources and references.}
\label{fgr:chip_architecture_overview}
\end{figure}

\clearpage
\subsection{Historical Progress in Sub-Efficiencies}
\label{sec:subeff_progress}

\begin{figure}[H]
\centering
\includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
\caption{\textbf{Historical improvements in LED sub-efficiencies.} Panels A1-A2 show data for key metrics used to calculate forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show data for internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Spectral data related to spectral sub-efficiency is shown in \cref{fig:phosphor_spectrum}. Additional context for the panels is provided below. Source: own synthesis of data from the list of sources provided in \cref{sec:sources}. For an illustration of the different chip architectures shown, compare \cref{sec:chip_architecture}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
\label{fgr:subeff}
\vspace{-20mm}
\end{figure}

@@

@@

\textbf{Additional Notes on \cref{fgr:chip_architecture_overview} and \cref{fgr:subeff}}

The data shown in \cref{fgr:subeff} and \cref{fig:phosphor_spectrum} was used to determine the performance of best-in-class devices across different sub-efficiencies. The results of these efficiency calculations are shown in Figures 3-5 in the main text. \cref{fgr:chip_architecture_overview} provides important context by illustrating the associated changes in chip architecture. This show the evolution of LEDs from classical chips with lateral current spreading to chip-scale package flip-chip architectures. We provide two examples:

Panel B2 of \cref{fgr:subeff} shows how light-extraction efficiency can be improved from $\sim40\%$ to $\sim95\%$ by switching from a planar Indium-Tin Oxide (ITO) surface to a patterned sapphire substrate (PSS). This is listed as a spillover in Figure 4 in the main text. Panel A in \cref{fgr:chip_architecture_overview} shows where on the chip these changes take place and provides typical examples of the relevant chip types.

Panel A1 of \cref{fgr:subeff} shows forward voltage was reduced (and therefore forward voltage efficiency increased) over time. This trend is observable both within the same chip architectures and between chip architectures. The most recent data points for 2018-2019 show how classical chips still have the highest forward voltage. Panels A-D in \cref{fgr:chip_architecture_overview} provides typical examples of the relevant chip types.

\clearpage
\subsection{Non-Phosphor-Related LED Innovations}

@@

\clearpage
\subsection{Phosphor-Related LED Innovations}
\label{sec:innovation_phosphor}

@@

\subsection{Spectral Data for Phosphors}

@@

\clearpage
\subsection{Full List of Sources for Figures}
\label{sec:sources}

@@

\clearpage
\section{Supplementary Methods}

\subsection{Literature Review Process}
\label{sec:litrev}

At the first stage of the comprehensive literature review described in section 7.2 in the main text, we conducted a search in specialized patent and publication databases for the most highly cited and/or relevant publications and patents containing information about innovations and overall historical progress in white LED lighting technology.

As several key reviews of technological progress in this field show, the dominant design of white LED lighting that emerged as a result of this progress and diffused at global scale in the lighting market is the GaN-based warm white phosphor-converted LED. See also \cref{sec:prev_lit} for a review of these publications. We therefore decided to focus our analysis exclusively on the evolution of this variation of white LED lighting technology as the only variation that has achieved global market success. Accordingly, we restricted our literature search only to publications focused on GaN-based LED lighting by excluding LEDs with other prominent semiconductor material compositions (i.e., GaAsP, AlGaInP, GaAlAs, ZnO, and perovskites), and on applications of LEDs in lighting by excluding other applications related to solar energy, lasers, or emittance in non-visible parts of the spectrum (e.g., ultraviolet LEDs).

The search was conducted through the \textit{Elsevier Scopus} database using the query provided in \cref{tab:scopus}. The exclusion filters placed additional restrictions on topical areas to focus the search outcomes only on the scientific and engineering basis of LED technology. The top 300 most highly cited papers were analyzed in the outcomes of this search both for their own content, and as an input to the second stage of the literature search described in section 7.2 of the main text that ensured the comprehensiveness of our findings: an iterative ‘snowballing’ search for backward citations in the publications identified at the first stage.

\begin{table}[h!]
\caption{Elsevier Scopus advanced search query.}
\begin{pseudocode}
TITLE-ABS-KEY(("light-emitting" OR "light emitting") W/2 diode*)
AND NOT
TITLE-ABS-KEY(solar OR violet OR laser)
AND NOT
TITLE-ABS-KEY(perovskite OR GaAsP OR AlGaInP OR GaAlAs OR ZnO)
AND NOT
TITLE-ABS-KEY(organic OR OLED* OR polymer*)
AND NOT
SUBJAREA(ARTS OR BUSI OR DECI OR ECON OR PSYC OR SOCI)
AND NOT
SUBJAREA(MEDI OR NURS OR VETE OR DENT OR HEAL OR MULT)
AND NOT
SUBJAREA(AGRI OR BIOC OR IMMU OR NEUR OR PHAR)
AND NOT
SUBJAREA(envi OR EART OR comp)
\end{pseudocode}
\label{tab:scopus}
\vspace{-4mm}
\end{table}

Additional search was conducted in patent databases \textit{The Lens} and \textit{Google Patents}. The search was restricted to a relatively broad CPC patent class H01L 33 (‘Semiconductor devices with at least one potential-jump barrier or surface barrier specially adapted for light emission […]’) and its constituent subclasses, focusing on most highly cited granted patents in this patent class and manually excluding those of them that were not relevant to white LEDs. In practice, this meant that the first-stage patent search was restricted to patents granted after 1990.

Finally, to complement the search in publications and patents, we also surveyed the archived websites of key LED manufacturers for any materials dedicated to their interpretation of LED history, history of their own products, and major product announcements.

@@

\subsection{Metrics for Quantifying Technological Progress in LEDs}

@@

\clearpage
\subsection{Device Performance Metrics}
\label{sec:device_performance_metrics}

@@

\clearpage
\subsection{Consumer Experience Metrics}
\label{sec:metrics_consumer}

@@

\subsection{Background of Interviewed Experts}

@@

\clearpage
\subsection{Cost Model Structure and Limitations}
\label{sec:costmodel_structure}

@@

\subsection{Yielded Cost}
\label{sec:cost_computation}

@@

\subsection{Contribution of Variables}
\label{sec:contribution_variables}

@@

\section{Supplementary Discussion}

\subsection{Cost Model Sensitivity Analysis}

@@

\clearpage
\subsection{Cost Data Comparison with Industry Projections}
\label{sec:cost_model_calibration}
